---
title: "Vietnam's Power Development Plan 8 (PDP8) Analysis"
author: "Vu Thuy Tien Trinh"
output: pdf_document
---

## Overview
This report analyzes Vietnam's Power Development Plan 8 (PDP8), focusing on changes in installed capacity by energy source between 2030 and 2050.

The analysis has two objectives:
1. Describe the planned capacity structure.
2. Visualize how Vietnam's power mix is expected to change over time.


### Headline Metrics

- Total Capacity 2030: 150 GW  
- Total Capacity 2050: 573 GW  
- Growth Factor: 3.8x

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(ggplot2)
library(tidyr)
library(dplyr)
library(knitr)
library(scales)
```

## 1. Descriptive Analysis

```{r data_clean, include=FALSE}
pdp8_data <- read.csv("vietnam_pdp8_data.csv") %>%
  select(Energy_Source, Capacity_2030, Capacity_2050)

total_2030 <- sum(pdp8_data$Capacity_2030)
total_2050 <- sum(pdp8_data$Capacity_2050)

analysis_table <- pdp8_data %>%
  mutate(
    share_2030 = Capacity_2030 / total_2030 * 100,
    share_2050 = Capacity_2050 / total_2050 * 100
  )
```

```{r table1, echo=FALSE}
knitr::kable(
  analysis_table,
  digits = 2,
  booktabs = TRUE,
  caption = "Installed Capacity and Shares under Vietnam's PDP8",
  align = "lrrrr"
)
```

The dataset contains planned installed electricity generation capacity for six major energy source categories in 2030 and 2050.

## 2. Data Visualization

```{r plot_data, fig.width=11, fig.height=7, fig.align='center'}
plot_data <- analysis_table %>%
  select(Energy_Source, Capacity_2030, Capacity_2050) %>%
  pivot_longer(
    cols = starts_with("Capacity"),
    names_to = "Year",
    values_to = "MW"
  ) %>%
  mutate(Year = gsub("Capacity_", "", Year))

ggplot(plot_data, aes(x = Year, y = MW, fill = Energy_Source)) +
  geom_col(width = 0.6) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Vietnam's Power Mix Transformation (PDP8)",
    subtitle = "Installed Capacity Growth: 2030 vs 2050",
    x = "Year",
    y = "Capacity (MW)",
    fill = "Energy Source"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 11),
    legend.position = "right"
  )
```

## 3. Key Findings

- Total installed capacity increases substantially between 2030 and 2050.
- Coal declines to zero by 2050 under the PDP8 pathway.
- Solar and wind become major pillars of future electricity supply.
- The results suggest a structural transition toward a lower-carbon power system.